---
title: "Running the model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  bib/mapspam.bib
bibliography: bib/mapspam.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`run_spam` runs the selected model (`min_entropy` or `max_score`) in GAMS. If `model_sel = 1`, the model is run for each individual administrative unit at level 1. If `model_sel = 0` the model is run only once for the total country. Selecting `out = TRUE` (default setting), the model log will be sent to the screen after the model run is finished. The log is as text file, whicn names starts with `model_log_` and is saved in the `processed_data/intermediate_output` folder.  Note that, depending on the size of the country and the selected resolution, the model might take a lot of time to run. If the model is very large, there is a risk your computer runs out of memory and an error message will be printed in the model log. 

```{r eval = FALSE}
run_spam(param, )
```

SPAMc results are saved into one (`solve_level = 0`)  or multiple (`solve_level = 1`) gdx files after the `run_spam` is executed. `combine_results()` combines the file(s) into a single rds file, that can be easily loaded into R and used for further inspection and further processing. The file is saved in the `processed_data/results` folder.

```{r eval = FALSE}
combine_results(param)

```


